---
title: "nycflights13"
author: "Author : Thunpischa Yodkaew"
date: "2022-12-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## NYCFlights13 Notebook

### load library

First step is loading library

-   `nycflights13` contains dataset about flights in NYC in 2013

-   `tidyverse` is used for manipulating data.

-   `lubridate` is used for managing datetime data.

```{r}
library(nycflights13)
library(tidyverse)
library(lubridate)
```

### 1. Which 5 airlines have the most average departure delay (in minutes).

```{r cars}

flights %>% 
  filter(dep_delay > 0) %>%  # filter only positive delay
  group_by(carrier) %>% 
  summarise(mean_delay = mean(dep_delay)) %>% 
  arrange(desc(mean_delay)) %>% 
  left_join(airlines, by = "carrier") %>% 
  select(airline = name,
         code = carrier,
         mean_delay_mins = mean_delay ) %>% 
  head(5)

```

SkyWest, Mesa, ExpressJet, Endeavor, Frontier Airlines are top5 of deley departure in 2013, respectively.

### 2. The causes that make flight delay.
